---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

```{r polldat}

# once file has been downloaded use this
responses <- read.csv("https://github.com/sgaichas/HSpresentations/raw/refs/heads/main/Projects/MAFMC_Indicators/Priority%20Management%20Decisions%20(Responses).csv",header=T)

# separate question number and element name? numbers are duplicates so just use name

#names(responses) <- sub("^X.*\\.\\.\\.","",names(responses))
names(responses)[-1] <- c("Priority1", "Scope1", "Species1", "Priority2", "Scope2", "Species2")

lookup <- c(
  #long = short
  "Develop a detailed example tailored to a particular species (Please write in species below)" = "Detailed example",
  "Develop a general process that could work across species/FMPs" = "General process"
)

responses$Scope1Brief <- unname(lookup[responses$Scope1])
responses$Scope2Brief <- unname(lookup[responses$Scope2])

respsumm <- responses |>
  dplyr::select(Priority1, Scope1Brief, Species1, Priority2, Scope2Brief, Species2) |>
  tibble::rownames_to_column("ID") |>
  tidyr::pivot_longer(-ID, names_to = "Var", values_to = "Val") 
  
totprior <- respsumm |>
  dplyr::filter(Var %in% c("Priority1", "Priority2")) |>
  dplyr::group_by(Val) |>
  dplyr::summarise(n = dplyr::n(),
                   perc = n/nrow(responses)*100) |>
  dplyr::arrange(desc(perc))

prior1 <- totprior[1,]

prior2 <- totprior[2,]

```

# MAFMC Indicators Project: Mini-Workshop Results

Workshop objective: Outline development of indicators and processes to use them in 2 priority Council management decisions.

## Mini-Workshop Introduction 

::: {.noteblock data-latex=""}
[Presentation](https://sgaichas.github.io/HSpresentations/20260331_ProjectOverview_MiniWK_Gaichas.html):  

1. A quick overview of the [project](https://www.mafmc.org/s/Project-5_IRA-Project-Brief.pdf) and [decisions made at Workshop 1](https://www.mafmc.org/s/WK1_Results-Summary.pdf)  

2. Reviewed pre-workshop poll: prioritized [Risk Policy](https://www.mafmc.org/s/3_MiniWK_RiskPolicy.pdf), and [Catch/Quota Setting](https://www.mafmc.org/s/4_MiniWK_CatchQuotas.pdf) 

3. Reviewed current processes and information available for Risk Policy and Catch/Quota Setting
:::

::: {.ltgreenbox data-latex=""}
## Workshop Discussions and Outcomes 

**Goal:** Prioritize 2 decisions and associated indicators for further development and testing of targets and thresholds 

**Discussion 1:** Discussed the relative merits of developing indicators for each priority management decision 

**Discussion 2:** Identifed desirable and undesirable states for the priority management outcomes 

**Discussion 3:** Discussed potential targets or thresholds associated with indicators related to management decisions
:::

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
::: {.ltpurplebox data-latex=""}
**Prioritization Poll Results** 
```{r poll, ft.arraystretch = 1}
#latex ignores flextable::line_spacing so ft.arraystretch argument in chunk header is as close as we get

flextable::flextable(totprior) |>
  flextable::set_header_labels(Val = "Decision",
                               n = "Votes",
                               perc = "% Total Votes") |>
  flextable::colformat_double(digits = 0) |>
  flextable::width(width = c(1.5,0.5,0.8)) |>
  flextable::fontsize(size = 9, part = "all") 
```
**Scope:** All who prioritized "Risk Policy" would focus on "General" scope. 
"Catch/Quota Setting" had 57% recommending a "Detailed example" for one species.
:::
## Discussion 1
*  Confirm management objectives for two priority decisions 
*  Do the decisions already have indicators available or information in hand?
*  Should we examine linked decisions independently?


## Discussion 2 
*  What is a desirable outcome for the management decision? 
*  What is undesirable?
*  A **good decision** is responsive and timely, incorporating new information as it becomes available, providing more flexible opportunities for fishing, and 
includes a clear description of tradeoffs


## Discussion 3
*  Indicator targets are points or ranges associated with desirable states
*  Indicator thresholds are points transitioning to undesirable states and require action if crossed
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}
*Project Steps: This mini-workshop focused on Management Decisions and linked Objectives, Information Supporting Decisions, links from Objectives to Targets and Thresholds, and preliminary Indicator Development*

```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject-miniwk.png")
```

::: {.ltbluebox data-latex=""}
*This workshop addressed project goals:*  
*  Review and assess existing indicators and evaluate their utility to support management objectives  
*  Identify and develop priority ecosystem and habitat indicators, including associated targets and thresholds as appropriate, and identify specific pathways for incorporation into Council management processes 
:::
:::
::::::



