---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

# MAFMC Indicators Project: Mini-Workshop Overview

Workshop objective: Scope development of operational ecosystem indicators for 2 priority Council management decisions.

## Mini-Workshop Introduction 

::: {.noteblock data-latex=""}
Presentations:  

1. A quick overview of the [project](https://www.mafmc.org/s/Project-5_IRA-Project-Brief.pdf) and [decisions made at Workshop 1](https://www.mafmc.org/s/WK1_Results-Summary.pdf)  

2. Review pre-workshop poll results: which 2 decisions are highest priority? What scope is preferable? 

3. Review SMART rating and information available for the 2 priority decisions
:::

::: {.ltgreenbox data-latex=""}
## Workshop Discussions and Desired Outcomes 

**Goal:** Prioritize 2 decisions and associated indicators for further development and testing of targets and thresholds 

**Discussion 1:** Discuss the relative merits of developing indicators for each priority management decision 

**Discussion 2:** Identify desirable and undesirable states for the priority management outcomes 

**Discussion 3:** Discuss potential targets or thresholds associated with indicators related to management decisions
:::


:::::: {.cols data-latex=""}

::: {.col data-latex="{0.49\textwidth}"}
*Things to think about...*

## Discussion 1
*  Confirm management objectives for priority decisions 
*  Do the decisions already have indicators available or information in hand?
*  Should we examine linked decisions independently?


## Discussion 2 
*  What is a desirable outcome for the management decision? 
*  What is undesirable?
*  A **good decision** is responsive and timely, incorporating new information as it becomes available, providing more flexible opportunities for fishing, and 
includes a clear description of tradeoffs


::: {.ltpurplebox data-latex=""}
*Example: Measures setting* 

*  Desirable: range of opportunities is maintained or expanded across species fleets while applying necessary measures to an individual species 
*  Undesirable: applying necessary measures to an individual species reduces range of opportunities, resulting in excess pressure on other species/resources or lack of opportunities to keep fisheries viable 
:::

## Discussion 3
*  Indicator targets are points or ranges associated with desirable states
*  Indicator thresholds are points transitioning to undesirable states and require action if crossed

::: {.ltpurplebox data-latex=""}
*Example: Measures setting* 

*  Target: maintain 5 different species/area/gear opportunities each month 
*  Threshold: opportunities in a species/area/gear are 2 or fewer in a month 
:::
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}

*Project Steps: This mini-workshop focuses on Management Decisions and linked Objectives, Information Supporting Decisions, links from Objectives to Targets and Thresholds, and preliminary Indicator Development*

```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject-miniwk.png")
```

::: {.ltbluebox data-latex=""}
*This workshop addresses project goals:*  
*  Review and assess existing indicators and evaluate their utility to support management objectives  
*  Identify and develop priority ecosystem and habitat indicators, including associated targets and thresholds as appropriate, and identify specific pathways for incorporation into Council management processes 
:::
:::

::::::



