---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

```{r polldat}

# once file has been downloaded use this
responses <- read.csv("https://github.com/sgaichas/HSpresentations/raw/refs/heads/main/Projects/MAFMC_Indicators/Priority%20Management%20Decisions%20(Responses).csv",header=T)

# separate question number and element name? numbers are duplicates so just use name

#names(responses) <- sub("^X.*\\.\\.\\.","",names(responses))
names(responses)[-1] <- c("Priority1", "Scope1", "Species1", "Priority2", "Scope2", "Species2")

lookup <- c(
  #long = short
  "Develop a detailed example tailored to a particular species (Please write in species below)" = "Detailed example",
  "Develop a general process that could work across species/FMPs" = "General process"
)

responses$Scope1Brief <- unname(lookup[responses$Scope1])
responses$Scope2Brief <- unname(lookup[responses$Scope2])

respsumm <- responses |>
  dplyr::select(Priority1, Scope1Brief, Species1, Priority2, Scope2Brief, Species2) |>
  tibble::rownames_to_column("ID") |>
  tidyr::pivot_longer(-ID, names_to = "Var", values_to = "Val") 
  
totprior <- respsumm |>
  dplyr::filter(Var %in% c("Priority1", "Priority2")) |>
  dplyr::group_by(Val) |>
  dplyr::summarise(n = dplyr::n(),
                   perc = n/nrow(responses)*100) |>
  dplyr::arrange(desc(perc))

prior1 <- totprior[1,]

prior2 <- totprior[2,]

```

# MAFMC Indicators Project: Mini-Workshop Results

Workshop objective: Outline development of indicators and processes to use them in 2 priority Council management decisions.

## Mini-Workshop Introductory [Presentation](https://sgaichas.github.io/HSpresentations/20260331_ProjectOverview_MiniWK_Gaichas.html) 

::: {.noteblock data-latex=""}
Overview of the [project](https://www.mafmc.org/s/Project-5_IRA-Project-Brief.pdf) and [decisions made at Workshop 1](https://www.mafmc.org/s/WK1_Results-Summary.pdf)  

Reviewed pre-workshop poll: prioritized [Risk Policy](https://www.mafmc.org/s/3_MiniWK_RiskPolicy.pdf), and [Catch/Quota Setting](https://www.mafmc.org/s/4_MiniWK_CatchQuotas.pdf) 

Reviewed current processes and information available for Risk Policy and Catch/Quota Setting
:::

::: {.ltgreenbox data-latex=""}
## Workshop Discussions and Outcomes 

**Discussion 1:** Discussed the relative merits of developing indicators for each priority management decision 

**Discussion 2:** Identifed desirable and undesirable states for the priority management outcomes 

**Discussion 3:** Discussed potential targets or thresholds associated with indicators related to management decisions
:::

:::::: {.cols data-latex=""}
::: {.col data-latex="{0.49\textwidth}"}
::: {.ltpurplebox data-latex=""}
**Poll Results** 
```{r poll, ft.arraystretch = 1}
#latex ignores flextable::line_spacing so ft.arraystretch argument in chunk header is as close as we get

flextable::flextable(totprior) |>
  flextable::set_header_labels(Val = "Decision",
                               n = "Votes",
                               perc = "% Total Votes") |>
  flextable::colformat_double(digits = 0) |>
  flextable::width(width = c(1.5,0.5,0.8)) |>
  flextable::fontsize(size = 9, part = "all") 
```
<!--**Scope:** All who prioritized "Risk Policy" would focus on "General" scope. 
"Catch/Quota Setting" had 57% recommending a "Detailed example" for one species.-->
:::

## Discussion 1: Indicators and approach
*  **Risk policy:**  consider social or economic risk indicators in addition to stock status while retaining overall simplicty of the approach. Start by reviewing cases where the Council has suspended the risk policy. Retain broad scope for application across species.
*  **Catch/quota setting:** review and communicate information available/used throughout the process to ensure indicators are not "double counted". Consider inforamtion used in prioritization and assessment but focus on developing new indicators for SSC ABC, Monitoring Committee, and Council processes. Consider a detailed application for a particular species.

## Discussion 2: Desirable and undesirable states
*  **Desirable:**  transparent and relatively simple decision processes, decisions reflect reality, interannual catch stability, ability to assess tradeoffs across objectives
*  **Undesirable:** overly complex processes, cascading unforseen impacts of decisions resulting in reactive rather than proactive management, accounting for conditions in more than one place 

## Discussion 3: Targets and thresholds
*  **Characteristics:**  targets as ranges rather than points, surveillance indicators to determine if conditions leave target range, historical baselines could provide initial target ranges but recognize and adapt to changing conditions, apply targets similarly across fishery sectors
*  **Other considerations:** indicators that imply certainty or reduce uncertainty are desirable 

:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}
*Project Steps: This mini-workshop focused on Management Decisions and linked Objectives, Information Supporting Decisions, links from Objectives to Targets and Thresholds, and preliminary Indicator Development*

```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject-miniwk.png")
```

::: {.ltbluebox data-latex=""}
*This workshop addressed project goals:*  
*  Review and assess existing indicators and evaluate their utility to support management objectives  
*  Identify and develop priority ecosystem and habitat indicators, including associated targets and thresholds as appropriate, and identify specific pathways for incorporation into Council management processes  
*  Develop operational capability by including students and early career scientists
:::
:::
::::::

<!--
The next steps for the project are:

1. to evaluate information sources and develop indicators that address the priority management decisions,  
1. to develop and apply frameworks to test the use of indicators in each decision, and
1. to prepare interim results for a workshop outlining performance metrics in late 2026.
-->

