---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE,
                      dev = "cairo_pdf")

library(patchwork)

flextable::set_flextable_defaults(
  font.size = 9,
  padding.bottom = 0,  
  padding.top = 1, 
  line_spacing = 1,
  font.family = "Palatino")

# can I set up 2 columns with highlight boxes in each? yes

```

# Catch/Quota Setting Step 4: ABC Setting

::: {.noteblock data-latex=""}
Catch and Quota Setting is a multistep process involving decisions by the Northeast Regional Coordinating Committee (NRCC; including the Mid Atlantic and New England Councils, Atlantic States Marine Fisheries Commission, and NOAA's Greater Atlantic Regional Fisheries Office and Northeast Fisheries Science Center), assessment scientists, assessment peer reviewers, the Council's Scientific and Statistical Committee (SSC), and the Council. Information for each stock and fishery is reviewed and synthesized to determine catch and quota levels that meet management objectives.
:::

*  **Key question: What additional indicators could inform SSC decisions at the ABC Determination step in this process? Could habitat, ecosystem, or socioeconomic indicators improve the process?** 
*  **Scope: Should we develop a general process for using indicators in the ABC Determination step of the catch/quota setting process, or focus on a specific application for a single species? Which species?** 


:::::: {.cols data-latex=""}

::: {.col data-latex="{0.49\textwidth}"}
::: {.ltbluebox data-latex=""}
## What information is used now?


::: {.ltpurplebox data-latex=""}
**4. Determine Acceptable Biological Catch (ABC)**

The SSC uses reviewed assessment outputs within a structured framework to determine ABC from assessment uncertainty based on input data, model diagnostics and performance, and ecosystem considerations.
::: 

The MT assessment gives current status and short term projections of overfishing limit (OFL) that may or may not already include ecosystem information. SSC may discuss performance of short term projections or terminal year estimates in the assessment; ecosystem information can inform decisions about projections or terminal year estimates. The SSC may request new OFL estimates/projections from the assessment lead based on this discussion. 

\vspace{1ex}

The SSC uses its [OFL CV process](https://www.mafmc.org/s/Final-OFL-CV-guidance-document_06_24.pdf) to evaluate uncertainty arising from data quality, model selection/identification, retrospective analysis, empirical comparisons,  ecosystem factors, and recruitment stanzas. Ecosystem information can feed into: 

*  Ecosystem factors: currently evaluate [Climate Vulnerability Analysis](https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0146756) results for each stock, can include multiple other environmentally driven changes in the stock
*  Recruitment stanzas: links to environmental drivers can decrease recruitment uncertainty

Ecosystem information that is presented along with the assessment is easiest to use in the OFL CV determination, but nothing limits the SSC from using ecosystem information presented from another source. The SSC has based decisions on published literature in the past.

\vspace{1ex}

The OFL CV is used along with the estimate of current stock status in the MAFMC control rule to determine the ABC that is recommended to the Council
SSC often provides annually averaged constant ABC for a specified period (2-3 years) as well as annual ABC for each year at the request of the Council.




:::

:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}
::: {.ltgreenbox data-latex=""}
## Information sources for indicators 

[State of the Ecosystem](https://www.fisheries.noaa.gov/new-england-mid-atlantic/ecosystems/state-ecosystem-reports-northeast-us-shelf), [Risk Assessment](https://www.mafmc.org/s/2_MAB_RiskAssess_2024.pdf), and [Ecosystem and Socioeconomic Profiles](https://www.fisheries.noaa.gov/new-england-mid-atlantic/ecosystems/ecosystem-and-socioeconomic-profile-development-and-reports) conceptual models and indicators for MAFMC stocks could inform SSC discussion of terminal year estimates, short term projection performance, and assessment uncertainty. 

\vspace{1ex}

Similar indicators as above plus habitat indicators [EFH Habitat Mapper](https://www.fisheries.noaa.gov/resource/map/essential-fish-habitat-mapper), [NRHA Data Explorer](https://nrha.shinyapps.io/dataexplorer/) could further inform OFL CV discussions. Systematic translation of stock specific indicators into risk criteria as in the Pacific could simplify OFL CV discussions.

```{r}
Tab4 <- readr::read_csv("https://github.com/sgaichas/HSpresentations/raw/main/Conferences/BevanUW_2026-02-19/data/WestCoastRiskFramework.csv")

flextable::flextable(Tab4[,1:2])|>
  flextable::set_caption("Pacific Pilot Risk Levels and Indicator Criteria") |>
  flextable::width(width = c(0.6,2.5)) |>
  #flextable::fontsize(size=9, part="all") |>
  #flextable::fit_to_width(max_width = 3.5) |>
  flextable::bg(i = c(1:3), bg=c("#00FF0050", "white", "#FF000050"))
  #flextable::bg(i = c(1:3), bg=c(scales::alpha("green", 0.5), "white", scales::alpha("red", 0.5)))

```




### Potential Pathways

*  Refine OFL CV criteria for considering integrated ecosystem conditions
*  Develop criteria for considering ecosystem impacts in short term projections


::: {.ltpurplebox data-latex=""}
## Information gaps 

Ecosystem and Socioeconomic Profiles are currently unavailable for summer flounder, scup, mackerel, butterfish, surfclam, ocean quahog, blueline tilefish, spiny dogfish, and monkfish. 

::: 
:::
:::

::::::




