---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

# MAFMC Indicators Project: Workshop 1 Overview

Workshop objective: Identify and prioritize Council management decisions that could be supported or improved by operational use of ecosystem indicators.

## Introduction 

::: {.noteblock data-latex=""}
*  A quick [project overview](https://www.mafmc.org/s/Project-5_IRA-Project-Brief.pdf) for background  

*  *Definitions* of terms so we are all on the same page  

*  Example indicator uses in management, followed by a preliminary list of example management decisions
:::

::: {.ltgreenbox data-latex=""}
## Discussions

What management objectives are associated with each of the decisions?

Which management decisions should this project focus on? Why?

Which of the gathered indicators look useful to address the priority management decisions?
:::


:::::: {.cols data-latex=""}

::: {.col data-latex="{0.49\textwidth}"}

Some text about the project
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}


```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject.png")
```
:::

::::::


