---
title: |
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

# MAFMC Indicators Project: Workshop 1 Overview

Workshop objective: Identify and prioritize Council management decisions that could be supported or improved by operational use of ecosystem indicators.

## Workshop Introduction 

::: {.noteblock data-latex=""}
Presentations:  

1. A quick [project overview](https://www.mafmc.org/s/Project-5_IRA-Project-Brief.pdf) for background  

2. *Definitions* of terms so we are all on the same page  

3. Example indicator uses in management, followed by a preliminary list of example management decisions
:::

::: {.ltgreenbox data-latex=""}
## Workshop Discussions and Desired Outcomes 

**Goal:** Which management decisions should this project focus on? Why? 

**Discussion 1:** Review the preliminary list of decisions 

**Discussion 2:** Identify priority decisions 

**Discussion 3:** Identify key information for priority decisions
:::


:::::: {.cols data-latex=""}

::: {.col data-latex="{0.49\textwidth}"}
*Things to think about...*

## Discussion 1
*  Are there Council decisions that should be added to the list of those presented?  
*  Which decisions are unclear to you?   
*  Can anything be removed?  
*  *Key Question: What management objectives are associated with each of the decisions?*

## Discussion 2
*  Which management decisions most often require additional information?  
*  What would a better decision look like? e.g., Reduce uncertainty? Increase consideration of tradeoffs? Prevent perverse incentives? Provide more flexibility or a wider range of options for participants?  
*  Given the objectives for each decision type, would ecosystem information help make better decisions?  
*  *Key Question: Which management decisions should this project focus on? Why?*

## Discussion 3
*  Which of the gathered indicators look useful to address the priority management decisions?   
*  Are there priority decision points that don’t appear to be supported by gathered indicators?  
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.49\textwidth}"}

*Project Steps: This workshop focuses on Management Objectives and Decisions, but notes linkages to other steps*

```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject.png")
```

::: {.ltbluebox data-latex=""}
*This workshop addresses project goals:*  
*  Review and assess existing indicators and evaluate their utility to support management objectives  
*  Identify and develop priority ecosystem and habitat indicators, including associated targets and thresholds as appropriate, and identify specific pathways for incorporation into Council management processes 
:::
:::

::::::



