---
title:
output:
  pdf_document:
    keep_tex: true
    includes:
      in_header: latex/boxescolumns.tex
  word_document: default
  html_document:
    css: columns.css
    df_print: paged
geometry: margin = 0.5in
link-citations: true
urlcolor: blue
---

\vspace{-1cm}
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      warning = FALSE,
                      message = FALSE)

library(patchwork)

# can I set up 2 columns with highlight boxes in each? yes

```

# MAFMC Indicators Project: Workshop 1 Results

Workshop objective: Identify and prioritize Council management decisions that could be supported or improved by operational use of ecosystem indicators. Workshop materials: https://www.mafmc.org/council-events/2025/dec-10/indicators-workshop

## December 10, 2025 Workshop Attendance: 40 total

::: {.noteblock data-latex=""}
*In person:* 13 participants from the Council, Council Advisory Panels, Scientific and Statistical Committee (SSC), Council staff, and Atlantic States Marine Fisheries Commission staff, plus 4 from the project team. 

*Online:* 21 participants from the Council, Council Advisory Panels, SSC, Council staff, Northeast Fisheries Science Center staff, Project Oversight Team, and other observers, plus 2 from the project team. 
:::

::: {.ltgreenbox data-latex=""}
## Overview of Workshop Discussions and Outcomes 

**Goal:** Which management decisions should this project focus on? Why? 

**Discussion 1:** Reviewed, clarified, and updated the [initial list of management decisions](https://www.mafmc.org/s/3_WK1_MgtDecisions.pdf) 

**Discussion 2:** Identified 4 priority decisions for indicator development by the project team 

**Discussion 3:** Provided initial feedback on [SMART indicator review](https://mafmc.github.io/SMARTindicatorsData/) documents and tools for refinement
:::


:::::: {.cols data-latex=""}

::: {.col data-latex="{0.60\textwidth}"}
## Discussion 1: Clarify Decisions
*  Frame management decisions to reflect both restriction and liberalization 
*  Consider distinctions between: 
   +  Pre-season and inseason decisions 
   +  Short term and longer term planning horizons 
   +  Routine and emergency contexts 
*  Consider tradeoffs between objectives in decisions 
*  Consider uncertainty: indicators should reduce it 
*  Consider stages in the process where information is most useful 
*  Additions:  Quota management, Decisions regarding emerging fisheries, MMPA and ESA objective interactions with EFH, Strategic planning and Annual Action Plan, Shared governance 

::: {.ltgreenbox data-latex=""}
## Discussion 2: Priority Decisions
*  **Risk policy:**  consider risk indicators in addition to stock status 
*  **Catch/quota setting:** communicate information available/used throughout the process. Consider Assessment, SSC ABC, Monitoring Committee, and Council processes 
*  **Setting measures:** consider multispecies and multifleet context in the specifications process. Objective: maximize opportunities across multiple management bodies. 
*  **Spatial management measures:** consider changing distributions of fish and fleets within the framework process  

#### For all selected decisions:
*  Evaluate performance of current management against stated objectives to aid indicator selection 
*  Evaluate impact of data availability or degradation on uncertainty and indicator performance 
:::

## Discussion 3: Indicator Tools
*  SMART Indicator framework and tools a promising approach 
*  Add information on state level surveys and other information used in Habitat Assessment
*  Design indicator explorer tools to bring information into many different Council products and processes  
:::

::: {.col data-latex="{0.02\textwidth}"}
\ 
<!-- an empty Div (with a white space), serving as
a column separator -->
:::

::: {.col data-latex="{0.38\textwidth}"}

*Project Steps: This workshop focused on Management Objectives and Decisions, and outlined additional steps*

```{r projectoutline, out.width="100%"}
magick::image_read("https://github.com/sgaichas/HSpresentations/raw/main/docs/images/indicatorsproject.png")
```

::: {.ltbluebox data-latex=""}
*This workshop addressed project goals:*  
*  Review and assess existing indicators and evaluate their utility to support management objectives  
*  Identify and develop priority ecosystem and habitat indicators, including associated targets and thresholds as appropriate, and identify specific pathways for incorporation into Council management processes 
:::
:::

::::::



